### 2014-09-27 Pero_Transcriptome.tex
starting revisions

The raw sequence reads corresponding to the four tissue types were error corrected using the software bless version 0.17 \cite{Heo:2014cb} using kmer=25, based on the developer's default recommendations. The error-corrected sequence reads were adapter and quality trimmed following recommendations from MacManes \cite{MacManes:2014io} and Mbandi \cite{Christoffels:2014bg}. Specifically, adapter sequence contamination and low quality nucleotides (defined as \textsc{Phred} \textless 2) were removed using the program Trimmomatic version 0.32  \cite{Lohse:2012fg}. Reads from each tissue were assembled using the Trinity version released 17 July 2014 \cite{Haas:2013jq}. We used flags to indicate the stranded nature of sequencing reads and set the maximum allowable physical distance between read pairs to 999nt. The assembly was conducted on a linux workstation with 64 cores and 512Gb RAM. To filter the raw sequence assembly, we downloaded \textit{Mus musculus} cDNA and ncRNA datasets from Ensembl (\url{ftp://ftp.ensembl.org/pub/release-75/fasta/mus_musculus/}) and the \textit{Peromyscus maniculatus} reference transcriptome from NCBI (\url{ftp://ftp.ncbi.nlm.nih.gov/genomes/Peromyscus_maniculatus_bairdii/RNA/}). We used a blastN (version 2.2.29+) procedure (default settings, evalue set to 10$^{-10}$) to identify contigs in the \textit{P. eremicus} dataset likely to be biological in origin. This procedure, when a reference dataset is available, retains more putative transcripts that a strategy employing expression-based filtering (remove if TMP \textless 1  \cite{MacManes:2012bu}) of the raw assembly. We then concatenated the filtered assemblies from each tissue into a single file and reduced redundancy using the software cd-hit-est version 4.6 \cite{Li:2006hr} using default settings, except that sequences were clustered based on 95\% sequence similarity. This multi-fasta file was used for all subsequent analyses, including annotation and mapping. \\

@@

We then estimated gene expression on each of these tissue-specific datasets, which allowed us to understand expression patterns in the multiple tissues (Pero.tissue.xprs, will be made available on Dryad, until then on Dropbox (\url{https://www.dropbox.com/sh/2jwzd8p6n6eluco/AAAO3nSdXb_u4wtQZRTwqW9ia?dl=0})). Specifically, we constructed a Venn diagram ({\hypertarget{Figure 1}{Figure 1}}) that allowed us to visualize the proportion of genes whose expression was limited to a single tissue and those whose expression was ubiquitous. 66324 transcripts are expressed on all tissue types, while 13086 are uniquely expressed in the kidney, 2222 in the testes, 3580 in the brain, and 2798 in the liver. The kidney appears to an outlier in the number of unique sequences, though this could be the result of the recovery of more lowly expressed transcripts that may be the result of deeper sequencing.  \\

>>>>

The raw sequence reads corresponding to the four tissue types were error corrected using the software bless version 0.17 \cite{Heo:2014cb} using kmer=25, based on the developer's default recommendations. The error-corrected sequence reads were adapter and quality trimmed following recommendations from MacManes \cite{MacManes:2014io} and Mbandi \cite{Christoffels:2014bg}. Specifically, adapter sequence contamination and low quality nucleotides (defined as \textsc{Phred} \textless 2) were removed using the program Trimmomatic version 0.32  \cite{Lohse:2012fg}. Reads from each tissue were assembled using the Trinity version released 17 July 2014 \cite{Haas:2013jq}. We used flags to indicate the stranded nature of sequencing reads and set the maximum allowable physical distance between read pairs to 999nt. We elected to assembly reads derived from a single deeply sequenced individual (Peer360, a male) to reduce polymorphism and thus the complexity of the de Bruijn graph, which has important implications for runtime, hardware requirements \cite{Lowe:2014iv,Pop:2009dp}, and assembly contiguity \cite{Vijay:2012gy and MacManes unpublished observations}. Individual tissues were assembled independently, as we hypothesize that tissue specific isoforms would be reconstructed with higher fidelity that if all tissues were assembled together.

The assembly was conducted on a linux workstation with 64 cores and 512Gb RAM. To filter the raw sequence assembly, we downloaded \textit{Mus musculus} cDNA and ncRNA datasets from Ensembl (\url{ftp://ftp.ensembl.org/pub/release-75/fasta/mus_musculus/}) and the \textit{Peromyscus maniculatus} reference transcriptome from NCBI (\url{ftp://ftp.ncbi.nlm.nih.gov/genomes/Peromyscus_maniculatus_bairdii/RNA/}). We used a blastN (version 2.2.29+) procedure (default settings, evalue set to 10$^{-10}$) to identify contigs in the \textit{P. eremicus} dataset likely to be biological in origin. This procedure, when a reference dataset is available, retains more putative transcripts that a strategy employing expression-based filtering (remove if TMP \textless 1  \cite{MacManes:2012bu}) of the raw assembly. We then concatenated the filtered assemblies from each tissue into a single file and reduced redundancy using the software cd-hit-est version 4.6 \cite{Li:2006hr} using default settings, except that sequences were clustered based on 95\% sequence similarity. This multi-fasta file was used for all subsequent analyses, including annotation and mapping. \\

@@

We then estimated gene expression on each of these tissue-specific datasets, which allowed us to understand expression patterns in the multiple tissues (Pero.tissue.xprs, will be made available on Dryad, until then on Dropbox (\url{https://www.dropbox.com/sh/2jwzd8p6n6eluco/AAAO3nSdXb_u4wtQZRTwqW9ia?dl=0})). Specifically, we constructed a Venn diagram ({\hypertarget{Figure 1}{Figure 1}}) that allowed us to visualize the proportion of genes whose expression was limited to a single tissue and those whose expression was ubiquitous. 66324 transcripts are expressed on all tissue types, while 13086 are uniquely expressed in the kidney, 2222 in the testes, 3580 in the brain, and 2798 in the liver. The kidney appears to an outlier in the number of unique sequences, though this could be the result of the recovery of more lowly expressed transcripts or isoforms.  \\
